---
title: "Regressions Final"
author: "JRP"
date: "12/30/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggpubr)
```

# Regression Models

## EXECUTIVE SUMMARY

In this quick analysis we will explore the relationship between a set of available variables to determine how to maximize miles per gallon fuel efficiency (MPG) based on the mtcars dataset. We will achieve this through a series of models that will help us determine the most effective way of predicting MPG, including the effects of having a manual or automatic transmission.

## AUTO / MANUAL ANALYSIS

The first question we seek to answer is whether an automatic or a manual transmission has better MPG. We can do this with a simple chart as observed in Fig 1 in the appendix. Based on the graph we can observe that manual transmissions have a clearly higher MPG. We can quantify this further with a quick linear model.

```{r a, echo = FALSE}
mtcars$trans <- ifelse(mtcars$am == 0,"auto", "manual")
fit <- lm(mpg ~ am, mtcars)
summary(fit)
fit$coefficients[1]

sum(fit$coefficients)

```

Based on our coefficients we can see that an automatic transmission has a mean predicted value of 17.15 MPG whereas a manual transmission has an estimate of 24.39 MPG. This can be observed in Fig 1 as well.

However, we get a low R-squared of 0.36. Even though "am" is statistically significant at p= 0.001 we can use the other variables to see if we get a better model to predict MPG.

## MODEL ANALYSIS

As shown in the first table in  the appendix, running a linear model including all other variables (aov) we can see that "cyl", "disp", and "wt" are the only statistically significant variables, with "am" failing even at p = 0.1. We next create a model that uses these variables, plus we'll retain "am" for now.

```{r b, echo = TRUE}
fit2 <- lm(mpg ~ am + cyl + disp + wt, mtcars)

```

Model 2's R-squared increases to 0.83, but it is likely overfitted and "am" is still not significant, nor is "disp". We try a final model and compare the results of all three.

```{r c, echo = TRUE}
fit3 <- lm(mpg ~ cyl+wt, mtcars)
```

The third model (fit3) that includes "cyl" and "wt" still has a rounded R-squared of 0.83, meaning we have lost very little explaining power by removing the other variables.

When we compare all three models in anova, we see that fit3 (mpg ~ cyl + wt) shows the greatest improvement and also has the lowest AIC, making it the best model.

Lastly, we do some graphical analysis in Fig 2- Fig 5, that shows a good distribution of residuals to fitted values, plus a normal distribution that we further confirm in the Shapiro-Wilk normality test.

We can also see in the expanded QQ plot in Fig 6 that the values remain within the confidence bands.

# APPENDIX

```{r d, echo = TRUE}

#Model Comparison and Statistical Tests

aov <- aov(mpg ~ cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb, mtcars)
summary(aov)

fit2 <- lm(mpg ~ am + cyl + disp + wt, mtcars)
summary(fit2)

fit3 <- lm(mpg ~ cyl+wt, mtcars)
summary(fit3)

anova(fit, fit3, fit2)

AIC(fit, fit2, fit3)

shapiro.test(fit3$residuals)
```

```{r pressure, echo=FALSE}

ggplot(data = mtcars, mapping = aes(x = trans, y = mpg))+
        geom_boxplot(aes(fill = am))+ theme(legend.position = "none")+
        xlab("Transmission")+ ggtitle("Fig 1. MPG Comparison between Transmissions")

par(mfrow = c(2,2))
plot(fit3, caption = c("Fig 2. Residuals vs. FItted",
                       "Fig 3. Normal Q - Q",
                       "Fig 4. Scale-Location",
                       " ", "Fig 5. Residuals vs. Leverage"))

ggqqplot(fit3$residuals, title = "Fig 6. QQ Plot showing Normality")

```


